
\subsection{Free Groups}

Free groups: Let $X$ be any set, consider the symbols of the form $x^n$ for $x \in X$, $n \in \bZ$. Consider all finite words in these symbols, ie. words of the form $x_1^{n_1} \cdots x_{k}^{n_k}$, with $n_j \in \bZ$ and a symbol $e_0$. If for some $j$, have $x_j = x_{j+1}$ then replace $x_j^{n_j}x_{j}^{n_{j+1}}$ by $x_j^{n_j + n_{j+1}}$, except if $n_j + n_{j+1}=0$, in which case omit $x_j^{n_j}x_{j}^{n_{j+1}}$ unless omitting gives the empty word, in which case replace it by $e_0$. \\
This procedure gives us ``reduced'' words. Denote the set of reduced words $G(X)$. Define a product on $G(X)$ by concatenation. $(x_1^{n_1} \cdots x_k^{n_k})(y_1^{m_1} \cdots y_l^{m_l}) = x_1^{n_1} \cdots x_k^{n_k}y_1^{m_1} \cdots y_l^{m_l}$, then reduce. 
\begin{itemize}
    \item Tedious computation shows that this product is associative. 
    \item Have inverses $(x_1^{n_1} \cdots x_k^{n_k})(x_k^{-n_k} \cdots x_1^{n_1})$ 
\end{itemize}
$G(X)$ is called the free group on the alphabet $X$. We get a map $\vp: X \to G(X)$ by $\vp(x) = x^1$ \\
Universal Property: If $H$ is any group, and if $f: X \to H$ is a function, there is a unique group homomorphism, $\hat{f}: G(X) \to H$ such that $\hat{f} \circ \vp = f$. 
\begin{itemize}
    \item Can be given explicitly, by $\hat{f}(x_1^{n_1} \cdots x_k^{n_k}) = f(x_1)^{n_1} \cdots f(x_k)^{n_k}$ 
\end{itemize}

\noindent
We get the following diagram and see that $G$ is a functor from the category of sets to the category of groups. 
\[\begin{tikzcd}
    X \arrow[r, "\phi^{X}"] \arrow[d, "h"] & G(X) \arrow[d, "G(h)"] \\ 
    Y \arrow[r, "\phi^{Y}"] & G(Y)
\end{tikzcd} \]

\begin{itemize}
    \item For each $x \in X$, $C_x = \{x^n \, : \, n \in \bZ\}$ is a subgroup of $G(X)$. The $C_x's$ for different $x \in x$ are independent so $G(X)$ is the free product of of the $C_x \cong \bZ$ 
\end{itemize}

\noindent
let $H$ be a group and let $S \subset H$ be a set of generators for $H$. We get $\hat{i}$ a group homomorphism onto $H$. Consdier $\ker(\hat{i})$, normal subgroup. Find a subset $R$ in $\ker(\hat{i})$ such that $\ker(\hat{i})$ is the smallest normal subgroup containing $R$, ie. the collection of all conjugates of elements of $R$ by elements of $G(S)$, then $G(S)/$ normal subgroup containing $R \cong H$, called presentation of $H$ by generators $S$ and relations $R$. \\

\noindent
$H \stackrel{h}{\to} A$ with $A$ commutative, then $h(xyx^{_1}y^{-1}) = e_A$, called commutator of $x$ and $y$, normal subgroup generated by commutators is in $\ker h$. Then if $H$, generators $S$, and relations contain all $sts^{-1}t^{-1}$ $s,t \in S$, then $G(S)/N(R)$ is a commutative group and any homomorphism of $H$ into an abelian group factors through $G(S)/N(R)$ 