
\subsection{Categories}

\begin{example}
    \begin{tabular}{c c}
        Objects & Morphisms \\
        Groups: $G$, homomorphimss $G \stackrel{f}{\to} H$ \\
        Sets: $S$, functions $S \stackrel{f}{\to} T$ \\
        Topological spaces: $X$, continuous maps $X \stackrel{f}{\to} Y$ 
    \end{tabular}
\end{example}

\noindent
Axioms 
\begin{itemize}
    \item For any two objects we have a set of morphisms $A \to B$ Mor($A,B$). 
    \item Can compose morphisms: $A \to B, B \to C$, we get a morphism $A \to C$ \\
    \item Identity morphism: $I_A: A \to A$ satisfying $f: A \to B$, $I_bf = f$< $fI_A = f$ 
    \item Function composition is associative
\end{itemize}

\begin{example}
    RIngs, varities, and differentiable manifolds
\end{example}

\begin{example}
    A group. Object: only 1 object. Morphisms: elements of group. Composition is group product. 
\end{example}

\begin{example}
    A poset (partially order set). Set $S$ with $\le$. Category: objects = elements of $S$. Morphisms: morhpisms from $A$ to $B$, 1- morphism if $A \le B$, none if $A \not\le B$. 
\end{example}

\noindent
Basic Theme: Ignore structure of objects, define everything using morphisms. \\
epimorphisms: analogs of surjective maps. Normal definition of surjective uses internal structure of $T$. \\
$f: S \to T$ is an epimorphism if whenever 2 morphisms $T \overset{g}{\underset{h}{\rightrightarrows}} U$ if $gf = hf \to g = h$ 

\begin{example}
    $f: S \to T$ ($S,T$ sets) $f$ surjective $\ifff f$ is an epimorphism 
\end{example}

\noindent
Warning: Sometimes epimorphism $\neq$ surjection 

\begin{example}
    Look at category of rings (morphisms = homomorphims) \\
    $f: \bZ \to \bQ$ not surjective but is an epimorphism of rings.  
\end{example}

\noindent
Fawcett: In category of planar graphs 4 color theorem $\ifff$ epimorphisms are surjective. 

\noindent
Dual Concept: Dual of surjectivity is injectivity \\
monomorphism: $f: S \to T$, if $R \overset{g}{\underset{h}{\rightrightarrows}} T$ $fg = gh \to g=h$ 

\begin{example}
    If $S, T$ subsets, $f: S \to T$ is injectivity $\ifff$ $f$ is monomorphism (also true for rings, groups, $\ldots$)
\end{example}

\subsection{Functors} 

Original Idea: Category of topological spaces $\to$ category of abelian groups \\

(Insert Figure) \\ 

\noindent
If $C, D$ categroies, a functor from $C$ to $D$ consist of 
\begin{enumerate}
    \item Object $F(X)$ for each object $X \in C$ 
    \item Morphism $f: X \to Y$ $\to$ morphism $F(f): F(X) \to F(Y)$ 
\end{enumerate}
Axioms: Behaves in ``obvious'' way. $F(\text{id}_A) = \text{id}_{F(A)}$, $F(fg) = F(f)F(g)$ 

\begin{example}
    Forgetful Functor, (Category of Groups) $\to$ (Category of Sets) by $G \mapsto$ underlying set, $G \to H \mapsto G \to H$ 
\end{example}