
\subsection{Initial Topologies}

\begin{definition}
    Let $X$ be a set. Suppose we have a collection of topologies $(Y_{\alpha}, \cT_{\alpha})$, and for each $\alpha$ a function $f_{\alpha}: X \to Y_{\alpha}$. The smallest topology $\cT$ such that each $f_{\alpha}$ is continuous is called the initial topology. 
\end{definition}

\noindent
For each $\alpha$, $U \in \cT_{\alpha}$, must have $f^{-1}_{\alpha}(U) \in \cT$ so a subbase of $\cT$ is $\{f^{-1}_{\alpha}(U) \, : \, \text{for all } \alpha, U \in \cT_{\alpha}\}$ 

\begin{example}
    Have $(Y, T^Y)$, let $X$ be a subset of $Y$. $f: X \hookrightarrow Y$ by $f(x)=x$. \\
    Inital topology on $X$ has subbase $f^{-1}(U) = U \cap X \subseteq X$ for $U \in \cT^Y$. Further, $\{ U \cap X \, : \, U \in \cT^Y \}$ is a topology. This topology is called the relative topology on $X$. 
\end{example}

\begin{example}
    $Y = \bR$, $X = [0,1]$, relative topology contains $[0, \frac{1}{2})$, not in the original topology 
\end{example}

\begin{example}
    Let $(X, \cT^X), (Y, \cT^Y) $ be topological spaces. Form set $X \times Y = \{(x, y) \, |\, x \in X, y \in Y \}$. \\
    We have projections $p^X: X \times Y \to X$ and $p^Y : X \times Y \to Y$. The initial topology has basis $(p^X)^{-1}(U) = U \times Y$ for $U \in \cT^X$, $(p^Y)^{-1}(V) = X \times V$ for $V \in \cT^Y$. \\
    Further, $(U \times Y) \cap (X \times V) = U \times V$ (rectangle) so the open sets of the initial topology consist of all arbitrary unions of rectangles $U \times V$ for $U \in \cT^X, V \in \cT^Y$, called the product topology on $X \times Y$.
\end{example}

\begin{example}
    $\bR^2 = \bR \times \bR$. The product topology contains rectangles $(a,b) \times (c,d)$\\
    Gives same topolgy as the euclidean metric 
\end{example}

\begin{itemize}
    \item Given $(X_1, \cT^{X_1}), (X_2, \cT^{X_2}), \ldots, (X_n, \cT^{X_n})$ can form $X_1 \times X_2 \times \cdots \times X_n$ with projections $p_1 : X_1 \times X_2 \times \cdots \times X_n \to X_i$. The product topology is generated by ``rectangles'' $U_1 \times U_2 \times \cdots \times U_n$ with $U_i \in \cT^{X_i}$ 
    \item Suppose for $n \in \bN$ we have $(X_n, \cT^n)$, can form $\Pi X_n$ with $p_j: \Pi X_n \to X_j$, $\fa j$. \\
    Only needs to contain finite intersections so we have a base of $U_1 \times U_2 \times \cdots \times U_m \times X_{m+1} \times X_{m+2} \times \cdots $ with $U_j \in \cT^j$. 
\end{itemize}

\begin{example}
    $X_j = \{0,1\}$ with discrete topology. $\Pi_{j=1}^{\infty} X_j$ not discrete, also compact. 
\end{example}

\begin{example}
    $C([0,1])$, set of continuous functions on $[0,1]$, $||f||_{\infty} = \sup \{f(t) \, : \, t \in [0,1]\} \to $ metric $d(f,g) = ||f-g||_{\infty}$ \\
    Given an normed vector space $(V, || \, ||)$, let $V'$= all continuous linear functionsal on $V$. \\
    eg. for $g \in C([0, 1])$ we have $\vp_g(f) = \int_0^1 f(t)g(t)dt$\\
    For $C([0,1]) \stackrel{\vp_g}{\to} \bR$, given topology not the smallest that makes each $\vp_g$ continuous. 
\end{example}